---
title: "Regression analysis"
author: "Berhe"
date: "2022-12-10"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      error = TRUE)

```


```{r load the required packages}
# load required packages
#Load in our packages - note the new packages we have added, emphasised by a blank line above and below them
pacman::p_load(
  rio,            # for importing data
  here,           # for relative file paths
  janitor,        # for cleaning data and simple tables
  skimr,          # inspecting the data
  lubridate,      # handling dates and times
  corrr,          # explore correlations
  gtsummary,      # easy publication-ready tables and regression
  skimr,          # quick breakdown of dataframes

  caret,          # for variable selection and carrying out statistical modelling and machine learning
  glmnet,         # needed to carry out variable selection
  lme4,           # for multivariable regression with random effects
  broom,          # for cleanly displaying multivariable regression results
  broom.mixed,    # also for cleanly displaying multivariable regression results
  tidyr,          # for cleaning results
  stringr,        # for pattern matching in trings

  tidyverse       # general data management and visualization
)

```


```{r import clean data}

data_clean<-import("data_clean.rds")

data_clean<-data_clean %>% 
  mutate(household_income=case_when(household_income=="0-4999"~"0-4999",
                                    household_income=="5000-9999"~"5000-9999",
                                    household_income=="10000-19999"~">= 20000",
                                    household_income=="20000-59,9000"~">= 20000",
                                    household_income==">60000"~">= 20000",
                                    TRUE~NA_character_)) %>%  
  mutate(household_income=fct_relevel(household_income,
                            "0-4999","5000-9999",">=20000")) %>% 
  mutate(age=case_when(age_in_years<18~"<18",
                                       age_in_years>=18 & age_in_years<23~"18-22",
                                       age_in_years>=23~"23+",
                                       TRUE~NA_character_)) %>% 
  
  mutate(delivery_choice_estimated_distance=case_when(delivery_choice_estimated_distance=="below 2 km"~"< 2 km",
                                            delivery_choice_estimated_distance=="2-5 km"~"2-5 km",
                                            delivery_choice_estimated_distance=="5-10 km"~">=5 km",
                                            delivery_choice_estimated_distance==">10 km"~">=5 km")) %>% 
mutate(listen_radio_programme=factor(listen_radio_programme,
                    levels = c("No",
                               "Yes",
                               NA_character_))) %>% 
mutate(watch_television=factor(watch_television,
                    levels = c("No",
                               "Yes",
                               NA_character_))) %>% 
mutate(ANC_during_last_pregnancy=factor(ANC_during_last_pregnancy,
                    levels = c("No",
                               "Yes",
                               NA_character_))) %>% 
mutate(residency_status=fct_relevel(residency_status, "rural","urban","unknown")) %>% 
mutate(attended_school=fct_relevel(attended_school, "No","Yes","unknown")) %>% 
mutate(ANC_time_to_first_visit=fct_relevel(ANC_time_to_first_visit,"months","weeks","unknown")) %>% 
mutate(Number_of_ANC_visits=fct_relevel(Number_of_ANC_visits,"1 visit",
                                        "2 visits", "3 visits","4+ visits", "unknown")) %>% 
mutate(time_travel_anc_visit=fct_relevel(time_travel_anc_visit,"0-30 minute","30-60 minute","60 plus minute")) %>% 
 mutate(delivery_choice_last_delivery =fct_relevel(delivery_choice_last_delivery, "health facility","home")) %>% 
  mutate(delivery_choice_traditional_medication=fct_relevel(delivery_choice_traditional_medication,"Yes","No")) %>% 
  mutate(delivery_choice_husband_choice=fct_relevel(delivery_choice_husband_choice,"home","health facility")) %>% 
  mutate(delivery_choice_estimated_distance=fct_relevel(delivery_choice_estimated_distance,"< 2 km","2-5 km",">= 2 km")) %>% 
  mutate(delivery_choice_time_taken=fct_relevel(
    delivery_choice_time_taken,"less than an hour","1-2 hours")) %>% 
  mutate(obstetrical_total_number_of_births=fct_relevel(obstetrical_total_number_of_births,"1","2-5",">5")) %>% 
  mutate(obstetrical_benfit_information=fct_relevel(obstetrical_benfit_information,"no","yes"))       


```

# Regression analysis
variables selected for the regression model 

```{r regression methods}
regression_data<-data_clean %>% 
  select(#age, 
         #age_group_1st_px,
         household_income,
         residency_status, 
         #marital_status,
         #respondent_religion,
         #respondent_occupation="occupation",
         attended_school,
         husband_education_level,
         #highest_educ_level_reached,
         husband_education_level,
         
         #husband_occupation,
         listen_radio_programme,
         watch_television,
# Antenatal care visit 
         ANC_during_last_pregnancy,
         ANC_time_to_first_visit,
         main_reason_for_attending_first_anc,
         #Number_of_ANC_visits,
         # ANC_means_of_transport_used,
         #time_to_travel_for_anc="time_travel_anc_visit",
        
# place of delivery for last delivery
         #place_of_delivery_last_baby="where_deliver_last_baby",
         #who_decided_last_delivery_place="delivery_decision",
         #means_of_transport_last_hf_delivery="meansof_transport_to_health_facility",
         time_taken_to_hf_during_last_delivery="distance_covered_tohealth_facility",
         services_satisfication_health_fcaility,
         #politely_treatment,
# choice for the last delivery
         delivery_choice_last_delivery,
         traditional_mediication_last_delivery="delivery_choice_traditional_medication",
         #last_hospital_delivery_assist="delivery_choice_assist_delivery_hospital",
         #last_delivery_deciesion_maker="delivery_choice_decision_fordelivery",
         husband_choice_for_place="delivery_choice_husband_choice",
         #road_acessible_for_delivery="delivery_choice_transport_accessibility",
         delivery_choice_estimated_distance,
         #transport_means_to_nearest_health_facility="delivery_choice_transport_means",
         time_to_reach_health_facaility_for_delivery="delivery_choice_time_taken",
         obstetrical_total_number_of_births,
         #obstetrical_age_at_first_pregnancy,
         obstetrical_benfit_information,
         #obstetrical_if_yes_how_many_visits_you_have_for_antenatal,
         overall_choice
          ) %>% 
     drop_na()

# check for contrast 
 #lapply(dataframe.df[c("x1", "x2", "x3")], unique)

```


## lets create cross tab for the variables to check the cell values before passing in to regression


```{r cross tab}

regression_data %>%
tbl_summary(by=overall_choice)

 
```



# Fit regression

 -LASSO regression fitting method was used as demonstrated below. it is the latest method of fitting logistic regression method as the stepwise method was found to have biases difficult to avaoid 

```{r then we create the test and train data set }

#This sets up fitting control

fitControl <- trainControl(
  method = "repeatedcv", 
  number = 5,         
  repeats = 10)   

#This runs the lasso regression
lasso_fit <- train(form = overall_choice ~ . ,
                   data = regression_data,    #This is the dataset that we created earlier
                   method = "glmnet",        #This is the name of the method that uses lasso regression
                   trControl = fitControl)   #This contains the fitting control arguments specified earlier

#Here we extract the predictor variables
lasso_predictors_raw <- predictors(lasso_fit)

#We can use the stringr function str_extract to clean the variable names
lasso_predictors_clean <- str_extract(string = lasso_predictors_raw,       #This extracts the important predictors from the Lasso fit
                                      pattern = paste(colnames(regression_data),   #This converts the column names into a REGEX expression, which allows us to extract all of these patterns from the predictors
                                                      collapse = "|"))

```

# Fit the multivariable qregression

In addition to the predictor variables selected using the LASSO predictors, if there are other variables that are known to be predictors by prior knowledge, add them in the final model manualy.

Hence, the below variables are added to the model in addition to the lasso predictors and one of the variables selected by the LASSO methods. Also the variables that are select by the LASSO and inflate the model prediction values are removed manualy. 

```{r fit multivariable refression}

# Multivariable regression ------------------------------------------------

#Now we have some variable suggestions, we can carry out our multivariable regression
#remember that the results of the lasso variable selection are NOT the only things you should include, add in other variables you think are important based on expert knowledge or important to your setting!
multivariable_regression_table <- regression_data %>%
  select(lasso_predictors_clean,
         -husband_choice_for_place,
         ANC_during_last_pregnancy,
         #ANC_time_to_first_visit,
         main_reason_for_attending_first_anc,
         #household_income,
         residency_status,
         #attended_school,
         husband_education_level,
         husband_education_level,
         obstetrical_total_number_of_births,
          overall_choice) %>% #Remember to include the outcome and any other variables not found in lasso_predictors_clean that you think are important!
  glm(formula = overall_choice ~.,                  #This formula structure uses all variables found in the dataset to explain outcome
      family = binomial) %>%                 #We are carrying out a logistic multivariable regression, so we want to use a binomial distribution (family)
  tbl_regression(exponentiate = TRUE) %>%    #As this is a logistic regression, we want the log odds and so we exponentiate
  add_nevent() %>% 
  bold_p()                                   #This highlights significant values
  
```

# Univariate regression 

add the the factors known by previous knowldege in addition to the LASSO predictor variables

```{r univariate regression analysis}
#We are now carrying out a uni-variate regression with the values used in the multivariable regression to compare them
univariate_regression_table <- regression_data %>%
  select(lasso_predictors_clean,
         -husband_choice_for_place,
         ANC_during_last_pregnancy,
         #ANC_time_to_first_visit,
         main_reason_for_attending_first_anc,
         #household_income,
         residency_status,
         #attended_school,
         husband_education_level,
         husband_education_level,
         obstetrical_total_number_of_births,
          overall_choice) %>%                         # select the previously defined variables
  tbl_uvregression(                                   # produce univariate table
    method = glm,                                     # define regression want to run (generalised linear model)
    y = overall_choice,                                      # define outcome variable
    method.args = list(family = binomial),            # define what type of glm want to run (logistic)
    exponentiate = TRUE                               # exponentiate to produce odds ratios (rather than log odds)
  ) %>%
  bold_p()
```

# Combine multivariate and univariate tables 

```{r merge the tables}

#Combine the tables
tbl_merge(tbls = list(univariate_regression_table, multivariable_regression_table),
          tab_spanner = c("Univariate", "Multivariable")) #This provides custom titles for the two regressions
```

# Forest plot 

```{r plot the forest plot}
multivariable_tidy <- multivariable_regression_table$model_obj %>% #To access the underlying data of the gtsummary object we can use $model_obj, this is what we need to generate our plot
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)  ## get a tidy dataframe of estimates

#Now we produce our forest plot
multivariable_tidy %>%
  subset(term != "(Intercept)") %>%    #We are removing the intercept from our plot as we are not interested in this
  ggplot(aes(x = estimate,             #This is our central estimate of the variable
             y = term,                 #This is the name of our variable
             xmin = conf.low,          #The minimum of our 95% confidence interval
             xmax = conf.high)) +         #The maximum of our 95% confidence interval
  geom_point() +                       #Plot a point at y
  geom_errorbar() +                    #Plot our errorbars at xmin and xmax
  geom_vline(xintercept = 1, linetype = "dashed") + #Extra step to add in a dashed line
  theme_bw() + #Extra steps for customising the output
  labs(x = "Estimate",
       y = "Term",
       title = "Odds ratio estimates") #Extra step to clean up the labels



# alternative


pacman::p_load("finalfit")
data(regression_data)
explanatory = c("time_taken_to_hf_during_last_delivery" , 
                "delivery_choice_last_delivery",              
                "delivery_choice_estimated_distance" ,        
                "time_to_reach_health_facaility_for_delivery", 
                "obstetrical_benfit_information",
                 "ANC_during_last_pregnancy",
                 #ANC_time_to_first_visit,
                 "main_reason_for_attending_first_anc",
                 #household_income,
                 "residency_status",
                 #attended_school,
                 "husband_education_level",
                 " husband_education_level",
                 "obstetrical_total_number_of_births" )
dependent = "overall_choice"

# Plot the the multivariate regression 
regression_data %>%
  or_plot(dependent, explanatory,
       table_text_size=2.2,
       title_text_size=5,
       plot_opts=list(xlab("OR, 95% CI"), 
       theme(axis.title = element_text(size=8))))

```


